Playing Games: A Case Study in Active Learning Applied to Game Theory

### 2015-06-30 paper/paper.tex
First draft complete
- Summary written
- Abstract written

Thorough read through required.

\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{subcaption}

@@

@@

\abstract{
A paper about active learning and using some example of this in a class on Game Theory
}

\section{Introduction}

@@

One of the inconsistancies of this approach is that all participants observe the
play by all the teams. Whilst from a mathematical perspective reputation is
inferred to mean the reputation gained during a particular duel, this has the
effect of teams being able to observe how other teams seem to play. A true
replication of Axelrod's tournament would not allow for this. One possibility
would invite participants to leave the room which might be logistically
constrained. From a pedagogic point of view however, having participants observe
the duels often leads to a much more engaged discussion (after as well as during
the activity).

@@

\section{Summary}

\begin{itemize}
    \item Give some examples of feedback.
    \item Mention how methods could be applied to other courses.
    \item Certain class management ideas (mainly that I will not speak first a
        lot of the time) <- Not sure if this is useful.
\end{itemize}

>>>>

\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{subcaption}

@@

% Diagrams
\usepackage{tikz}
\usetikzlibrary{calc}

@@

\abstract{
    This paper describes two active learning activities aiming to introduce
    students to the theoretic concepts of best response dynamics and repeated
    game analysis.

    An overview of some literature on active learning and the benefits therein
    is also provided. This highlights that activities such as the one described
    in this manuscript not only help engage students but more importantly
    improve their learning and understanding.

    The final section of the work describes how these activities fit in the
    pedagogic framework of a particular class. Students generate data that
    can be used as context for the understanding of theoretic concepts. It is
    suggested that this framework is not restricted to the subject of game
    theory.
}

\section{Introduction}\label{sec:introduction}

@@

One of the inconsistencies of this approach is that all participants observe the
play by all the teams. Whilst from a mathematical perspective reputation is
inferred to mean the reputation gained during a particular duel, this has the
effect of teams being able to observe how other teams seem to play. A true
replication of Axelrod's tournament would not allow for this. One possibility
would invite participants to leave the room which might be logistically
constrained. From a pedagogic point of view however, having participants observe
the duels often leads to a much more engaged discussion (after as well as during
the activity).

@@

\section{Summary and place within a pedagogic framework}\label{sec:summary}

These activities have been used by the author during outreach events during
which students take part in the activity of Section \ref{sec:best_responses} and
whilst the results of that are being analysed take part in the activity of
Section \ref{sec:repeated_games}. These two activities complement themselves and
form an accessible introduction to novel mathematical topics for a wide range of
age groups.

More notably however these activities have been used as part of a family of
activities used in a final year undergraduate course. This particular course is
taught in active learning pedagogic framework akin to a flipped class where
students are introduced to theoretic concepts through prior `playing of games'.
Other examples of these activities include:

\begin{itemize}
    \item \textbf{A rock paper scissors lizard tournament}: this serves as an
        introduction to mixed strategies.
    \item \textbf{A variety of games using coin flips}: this serves as an
        introduction to game with incomplete information.
    \item \textbf{Playing paper bin basketball in teams}: this serves as an
        introduction to cooperative game theory.
\end{itemize}

The general pedagogical basis for this is discussed in Section
\ref{sec:introduction} and the particular framework is shown in Figure
\ref{fig:use_of_data}. Students are active participants in the creation of
`data' which drives a discussion:

\begin{itemize}
    \item Why did you all guess this?
    \item Why did that team say that on that particular occasion?
    \item What would be a fair way of sharing the spoils for this particular
        game?
\end{itemize}

Following that discussion the theory can be put in context by highlighting
particular theoretic results and how they correspond (or not) to the behaviour
exhibited during the activity. Furthermore, this encourages an immediate
feedback with regards to student comprehension which can be reactively
addressed.

\begin{figure}[!hbtp]
    \begin{tikzpicture}
        % Nodes
        \node (activity) at (0,0) [fill=cyan!30, draw] {Activity};
        \node (data) at ($(activity) + (3,0)$) [fill=purple!30, draw] {Data};
        \node (discussion) at ($(data) + (2,1.5)$) [fill=green!30, draw] {Discussion};
        \node (theory) at ($(data) + (2,-1.5)$) [fill=red!30, draw] {Theory};

        % Creative arrows
        \draw [thick, ->] (activity) -- (data);
        \draw [thick, ->] (data) edge[out=90, in=180] (discussion);
        \draw [thick, ->] (discussion) edge[out=-45, in=45] (theory);

        % Reflective arrows
        \draw [thick, ->, dashed] (theory) edge[out=180, in=-90] (data);
        \draw [thick, ->, dashed] (theory) edge[out=180, in=-90] (activity);
        \draw [thick, ->, dashed] (discussion) edge[out=180, in=90] (activity);
        \draw [thick, ->, dashed] (theory) edge[out=135, in=-135] (discussion);

        % Key
        \draw [thick, ->] ($(activity) + (-3,.25)$) -- ($(activity) + (-2.5,.25)$);
        \node at ($(activity) + (-4, .25)$) {Production:};
        \draw [thick, ->, dashed] ($(activity) + (-3,-.25)$) -- ($(activity) +
        (-2.5,-.25)$);
        \node at ($(activity) + (-4, -.25)$) {Reflection:};

    \end{tikzpicture}
    \centering
    \caption{The active generation of data by students}
    \label{fig:use_of_data}
\end{figure}

This pedagogic approach is used throughout the course (from the first lesson)
and so after a few class meetings students are used to the high level of
participation. Here are some examples of written feedback concerning the
activities used in class.

\begin{itemize}
    \item Quote 1
    \item Quote 2
    \item Quote 3
\end{itemize}

Nonetheless at the start of the course certain class management
techniques described in \cite{Rocca2010} are used. For example, the extension of
the `waiting time' for responses to questions. For students to be active
participants it is vital that they are given the time to do so.

The activities described in Section \ref{sec:game_theory} are particular to game
theory however the author does not feel that the general pedagogic strategy
outlined in Figure \ref{fig:use_of_data} is constrained to a particular course.
Similar activities could be devised in other subjects where students generate
`data' than aids the contextualisation of theory so as to aspire to not only a
constructive learning model but also a social one.
